---
title: "Deepika Analysis"

---


```{r libraries}
library(ggplot2)
library(caret) 
library(glmnet)
library(MASS) 
library(e1071) 
library(mlbench) 
library(pROC) 
library(AppliedPredictiveModeling)
library(tidyverse)
```

```{r data import}
dat1<-read.csv("final.csv")
dat1$X <-NULL

#Data Partition
set.seed(1)
rowTrain <-createDataPartition(y=dat1$gender, p=0.8, list=FALSE)

#Set Gender as factor
dat1$gender <-as.factor(dat1$gender)
levels(dat1$gender) <-c("female", "male")

```


#Logistic regression: General
```{r}
#Cross Validation
ctrl <- trainControl(method = "repeatedcv", 
                     repeats = 5, 
                     summaryFunction = twoClassSummary, 
                     classProbs = TRUE) 

set.seed(1) 


log.fit <- train(x = dat1[rowTrain,1:19], 
                   y = dat1$gender[rowTrain], 
                   method = "glm", 
                   metric = "ROC", 
                   trControl = ctrl)
```


##Confusion Matrix: Logistic Regression
```{r}
test.pred <-predict(log.fit, newdata = dat1[-rowTrain,], type = "raw")

confusionMatrix(data = as.factor(test.pred), 
                reference = dat1$gender[-rowTrain], 
                positive = "male")

```


##ROC Curve: Logistic Regression
```{r}
test.pred.prob <-predict(log.fit, newdata = dat1[-rowTrain,], type = "prob")
roc.glm <- roc(dat1$gender[-rowTrain], test.pred.prob$male) 
plot(roc.glm, legacy.axes = TRUE, print.auc = TRUE) 
plot(smooth(roc.glm), col = 4, add = TRUE)


```

#Logistic Regression: Regularized
```{r}
#Tuning Gride
glmnGrid <- expand.grid(.alpha = seq(0, 1, length = 6), 
                        .lambda = exp(seq(-8, -2, length = 20))) 

set.seed(1) 
reg.fit <- train(x = dat1[rowTrain,1:19], 
                   y = dat1$gender[rowTrain],
                    method = "glmnet", 
                    tuneGrid = glmnGrid, 
                    metric = "ROC", trControl = ctrl)

plot(reg.fit, xTrans = function(x) log(x))

#The final values used for the model were alpha = 0 and lambda = 0.07196474
```

##Confusion Matrix: Logistic Regression (Regularized)
```{r}
test.pred2 <-predict(reg.fit, newdata = dat1[-rowTrain,], type = "raw")

confusionMatrix(data = as.factor(test.pred2), 
                reference = dat1$gender[-rowTrain], 
                positive = "male")

```

##ROC Curve: Logistic Regression (Regularized)
```{r}
test.pred.prob2 <-predict(reg.fit, newdata = dat1[-rowTrain,], type = "prob")
roc.glm2 <- roc(dat1$gender[-rowTrain], test.pred.prob2$male) 
plot(roc.glm2, legacy.axes = TRUE, print.auc = TRUE) 
plot(smooth(roc.glm2), col = 4, add = TRUE)

```

#Discriminant analysis (Linear)
```{r}
set.seed(1) 
model.lda <- train(x = dat1[rowTrain,1:19], 
                   y = dat1$gender[rowTrain], 
                   method = "lda", 
                   metric = "ROC", 
                   trControl = ctrl)

```


##Confusion Matrix: Discriminant analysis (Linear)
```{r}
lda.pred <- predict(model.lda, newdata = dat1[-rowTrain,], type = "raw")

confusionMatrix(data = as.factor(lda.pred), 
                reference = dat1$gender[-rowTrain], 
                positive = "male")
```

##ROC Curve: Discriminant analysis (Linear)
```{r}
lda.pred.prob <-predict(model.lda, newdata = dat1[-rowTrain,], type = "prob")
roc.lda <- roc(dat1$gender[-rowTrain], lda.pred.prob$male) 
plot(roc.lda, legacy.axes = TRUE, print.auc = TRUE) 
plot(smooth(roc.lda), col = 4, add = TRUE)

```

#Discriminant analysis (Quadratic)
```{r}
set.seed(1) 
model.qda <- train(x = dat1[rowTrain,1:19], 
                   y = dat1$gender[rowTrain], 
                   method = "qda", 
                   metric = "ROC", 
                   trControl = ctrl)

```


##Confusion Matrix: Discriminant analysis (Quadratic)
```{r}
qda.pred <- predict(model.qda, newdata = dat1[-rowTrain,], type = "raw")

confusionMatrix(data = as.factor(qda.pred), 
                reference = dat1$gender[-rowTrain], 
                positive = "male")

```

##ROC Curve: Discriminant analysis (Quadratic)
```{r}
qda.pred.prob <-predict(model.qda, newdata = dat1[-rowTrain,], type = "prob")
roc.qda <- roc(dat1$gender[-rowTrain], qda.pred.prob$male) 
plot(roc.qda, legacy.axes = TRUE, print.auc = TRUE) 
plot(smooth(roc.qda), col = 4, add = TRUE)

```
